Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach

### 2020-04-28 Response_to_PRD_reviewers.txt

Looking at Fig. 7, however, I question if it is possible to achieve
such low FAPs with this pipeline, particularly for the neutrino
mechanism waveforms. It would be useful to remake this figure at
distances for which FAP << 0.3% is possible. If not possible, it may
be beneficial to remove the figure. If not removed, the caption and
text should be updated to note the large FAPs used.

Response: We have tested again the trained CNN with 1.08^6 test samples 
          for each distance so that we can achieved a FAP of 0.1%, 0.01% 
          and 0.001%.

          We appreciate that re-running the simulations at closer
          distances is advised, but doing so would be very 
          computationally expensive.

          The changes in response to this comment includes: 
          1. The plots in the section of results (section IV and section V) 
             have been updated.  The text in these two sections have been 
             modified to reflect these updated results.

          2. Only TAP values at FAP equal to 0.1% are quoted in the      
             abstract and the conclusion. No detection claim is made for 
             FAP higher than 0.1%. The text in the abstract and the 
             conclusion has also been modified to accommodate the changes.

@@

          3. We retracted claims regarding the ability of the 
             networks in detecting Galactic events where the explosion 
             mechanism is the neutrino-driven mechanism.

          4. The number of training/test samples in the last paragraph of 
             section III has been updated to be consistent with 
             this change.

          5. We would also like to highlight the fact that the plots
             previously submitted as Fig. 9a and Fig. 9b were actually 
             duplicates of previous Fig. 6a and Fig. 6b by accident. 
             Although this mistake was naturally corrected when we updated
             the results, we would still like to bring this mistake to the
             reviewers' attention.

@@

Overall, the study is a novel and interesting work. It is crucial,
however, to put the authors' pipeline through more rigorous testing
(i.e. at realistic FAPs, perhaps using realistic noise) before stating
its ability to detect GWs from Galactic CCSNe. I would suggest framing
instead the paper as a follow-up pipeline, but take note that the
authors explicitly distinguish their pipeline from that shown in a
similar work by Astone+2018 as designed to take time series directly
from the detectors and find signals, rather than analyzing triggers
found by some external detection pipeline. Given this, it seems
appropriate for appreciable corrections to the text and results to be
requested.

Response: We appreciate this comment and admit that testing/training 
          the CNN using realistic noise would indeed show a more 
          realistic results.
          However, such an extension of the current work would require 
          a whole new set of treatments that we feel is beyond the current scope
          of this work. 
          We have therefore updated our results with more realistic 
          values of FAP and leave tests using realistic noise for future 
          work.

Below we list specific comments by section, denoting the
relevant paragraph number by p.

@@

p1: LIGO, Virgo, and KAGRA are not introduced or cited here. Suggest
to move discussion of second generation detectors (and references)
from start of p2 here. Aside - has KAGRA now begun operations? If so,
it may be worth updating the text here.

Response: We have updated the text accordingly.

p2: Specifically, the sentence "The nuclear equation of state will
then stiffens by the strong nuclear force above the nuclear density
and stops the core collapse." doesn't really make sense. The equation
of state stiffens (and subsequently stops the core collapsing) as the
core density exceeds nuclear density because the strong nuclear force
becomes repulsive at very small inter-nucleon distances.

Response: Thanks for pointing this out. We have rewritten the sentence 
          to read "The nuclear equation of state will then stiffens by 
          the strong nuclear force as the core exceeds nuclear density
          and stops the core collapse."

@@

Response: This has been corrected.

III: Data

p3: For the 3D simulations, what is the observer angle for the h+,hx
used? As the GW signal for CCSNe varies with orientation of the source
with respect to the observer, it may be useful to add a distribution
of observer angles for each waveform model to the training set, so as
to account for this variance. This could also help remedy the issue
noted in p6 of small sample size of neutrino mechanism waveforms
compared to magnetorotational mechanism waveforms.

Response: The observer angles for all the studies of 3D simulations have
          been added to the text. We appreciate that our data could be 
          augmented by varying the observer angles in the data set. 
          However, we noted that while some of the used studies published 
          the information or parameters to generate waveforms at random 
          observer angles, the same information for other studies is not
          readily available. We therefore decided to use only the published
          waveforms.

@@

Response: As stated above, we have updated section IV and section V.

V: Waveforms Excluded From Training

As for the results section, I suggest reruns using parameters
highlighted above, and the text subsequently rewritten to reflect the
updated results. Similarly to before, at the very least a disclaimer
on the FAPs must be added, and detection statements for FAP>0.3%
removed.

Response: As stated above, we again tested the trained CNN with 
          1.08 * 10^6 test samples to achieve FAP values of 0.1%, 0.01%, 
          0.001% as was done for section IV. The text and plots 
          are also updated to reflect this change. 
          Claims regarding the ability of the networks in detecting 
          Galactic CCSN signals if the explosion mechanism is the 
          neutrino driven mechanism are withdrawn.

### 2020-07-16 main.tex
Fixed plot style
revised paper 2nd

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae out to the Large or even Small Magellanic Clouds. 
%A Galactic event is likely to be detectable with a network of H+L+VK 
%if the explosion mechanism is the neutrino-driven mechanism.  
By testing the convolutional neural network 
with waveforms not used for training, we show that the true alarm probability
is $66\%$ at $60$ kpc for waveforms $\text{R3E1AC}$ and 
$\text{R4E1FC\_L}$ for networks of H+L+VK.  
For waveforms $\text{s}20$ and $\text{SFHx}$ at $10$ kpc, 
the true alarm probabilities are $49\%$ and $20\%$ for a network of H+L+VK 
and HLVK respectively. 
All at false alarm probability equal to $0.1\%$. 
\end{abstract}

@@

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to
\ac{aLIGO}, \ac{AdVirgo} and KAGRA interferometers. 
It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffens by the strong nuclear force as the core exceeds
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

@@

%
% introduce machine learning
% 
In recent years, the field of machine learning and its sub-field, deep
learning, have been rapidly developing and have shown great potential in many
scientific fields~\cite{krizhevsky2012imagenet, NIPS2014_5423,
simonyan2014very, chen2014semantic, zeiler2014visualizing, szegedy2015going}.
For example, deep learning has been successfully applied to fields including
medical diagnosis~\cite{kononenko2001machine}, object
detection~\cite{redmon2016you}, image
recognition/processing/generation~\cite{he2016deep, krizhevsky2012imagenet,
zhang2016colorful, karpathy2015deep}, and  language
processing~\cite{lample2016neural}. In \ac{GW} astronomy, deep learning has
mostly been applied to the identifications of detector noise artefacts
(glitches)~\cite{mukund2017transient, zevin2017gravity, george2017deep} and the
detection of astrophysical signals~\cite{george2018deep, gabbard2018matching, astone2018new},
and their parameter estimation~\cite{2019arXiv190906296G}.
An advantage of the use of a \ac{CNN} in the detection
of a \ac{GW} signal is that a \ac{CNN} is relatively computationally cheap
compared to other more traditional methods. This is because the heavy
computational work is usually done during the training stage of a \ac{CNN}
prior to its actual application~\cite{goodfellow2016deep}.

%
% the main point of the paper
%
In this work, we demonstrate that a \ac{CNN} can be applied to the detection of
\acp{GW} from \acp{CCSN} and to the classification of their explosion
mechanisms. 
To train our \ac{CNN}, we use simulated \ac{CCSN} waveforms from a
number of studies and simulate sources at a range of distances from $10$ to
$200$ kpc for two networks of four \ac{GW} detectors. The first network
consists of \ac{aLIGO}, \ac{AdVirgo} and KAGRA. For the second network, we
still include the detectors of \ac{AdVirgo} and KAGRA, but replace the two
detectors of \ac{aLIGO} with a modest set of planned upgrade version of them -
LIGO A+ in Hanford and Livingston~\cite{miller2015prospects, LIGOW}. All detectors 
are at their design sensitivities. We then
apply the \ac{CNN} to waveforms excluded from the training set to show 
the performance on an independent testing set.
Similar to~\cite{astone2018new}, 
we use machine learning techniques for the search of \acp{GW} from \acp{CCSN}.
However, the \ac{CNN} developed in ~\cite{astone2018new} takes time-frequency images of \ac{GW} detector 
data from the coherent WaveBurst pipeline~\cite{klimenko2005constraint} as input, 
while we aim to develop a \ac{CNN} that works independently and takes time series data from \ac{GW} detectors as input.

@@

%
% describe data pre-processing
%
Since the waveforms are generated at various distances, sampling rates and
durations, it is necessary to normalise the waveforms before they can be used
for the generation of the time-series. To do this, we first scale the
amplitudes of the waveforms by moving the sources to $10$ kpc from earth. We
then ensure that the sampling rate are identical for all the waveforms by down
sampling them to a pre-selected sampling rate of $4096$Hz. The longest duration
$\tau$ among the waveforms is then identified and each of the remaining
waveforms is padded with zeros to this duration. To introduce as few artefacts
as possible, a high pass filter with a low cut-off frequency equal to $11$Hz
and a tukey window ($\alpha = 0.08$) are applied prior to the zero padding.

%
% data augmentation
%
To balance the difference in the number of waveforms between the two mechanisms
in the training, validation and testing data set, $36$ exact copies of each neutrino-driven waveform are
made. This will make the ratio of the waveforms modelled by the two mechanisms close to 1, 
while does not change the ratio of the waveforms from different studies modelled by the neutrino-driven mechanism. 
This means each waveform in the neutrino-driven mechanism will have equal representations in the data set,
and waveforms from different mechanisms will have equal representation.
After this procedure, the simulated waveforms are then $\bold{S}(t) = \{\bold{s}_{1}(t), \bold{s}_{2}(t),
..., \bold{s}_{m}(t)\}$, where $m$ is the number of the waveforms and
$\bold{s}_{q}$ is the $q^{\text{th}}$ waveform, defined by, 
%
\begin{equation}\label{eq:ht}
\bold{s}_q(t) = \left(\begin{array}{c} h_q^+(t) \\ h_q^\times(t)
\end{array}\right), 
\end{equation} 
%
where $h_q^+(t)$ and $h_q^\times(t)$ are the two polarisations of the waveform
and $t$ is the time. In this work, we assume a 
fixed GPS time when generating the training/validation/testing data. 
At this GPS time, the value of $t$ is zero.
As mentioned, some of the waveforms used in the work are
generated with the simulations being axisymmetric, in which case the waveforms
are entirely described by one polarisation $h_q^+(t)$. The corresponding
$h_q^\times(t)$ for these waveforms are defined as a vector of zeros. 
%
% rescaling the amplitudes for differnet distances
%
In this work, we perform simulations for distances equal to $10$, $20$, $30$,
$40$, $50$, $60$, $80$, $100$, $150$ and $200$ kpc. If for a training
procedure, the default distance $d_\text{L}$ of a waveform is not $10$ kpc, the amplitude are rescaled 
by the distance since the amplitude is inversely proportional to the distance.

@@

The next step is to generate simulated time-series for the \ac{GW} detectors in

@@

%The input data consist of simulated time-series from a network of \ac{GW}

@@

into three exclusive classes, in total three types of time-series are

@@

\begin{equation}

@@

\bold{N}_j(T)~~~~~~~~~~~~~~~~~T < T_{\text{max}} p; \\

@@

\bold{N}_j(T)~~~~~~~~~~~~~~~~~T > \tau + T_{\text{max}} p, \\

@@

@@

For each time-series of the background noise class, independently simulated Gaussian

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 1.26 \times 10^{6}$ data samples, where each of the 3 classes
contains approximately $4.2\times10^{5}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1.0\times10^{4}$ samples being randomly selected for validation.
Of the remainder, $1.70\times10^{5}$ for training and $1.08\times10^{6}$ for tesing. 
The large number of testing samples is chosen so that low values of \ac{FAP} can be obtained 
(see section \ref{sec:result} and section \ref{sec:unseen}).
When the training is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

%
% plot of representative data sample
%
\begin{figure}
\includegraphics[width=0.475\textwidth]{datasample.png}
\caption{A representative example of the simulated time-series used to
train/validate/test the \ac{CNN}. The blue shows a whitened time-series with a
signal added to Gaussian noise generated using the power spectral density of \ac{aLIGO}. 
The red curve shows the same whitened signal free of
noise. The optimal signal to noise ratio for this signal is $5.5$.
The original waveform is shown in Fig. \ref{fig:neusample1}.
The source is at a distance of $10$ kpc from earth.
\label{fig:sample}} 
\end{figure}

@@

%
% describe ROC curve plots (Fig 5)
%
In Fig.~\ref{fig:ROClog}, we show the \acp{ROC} for both of the \ac{CCSN}
mechanisms and \ac{GW} detector networks tested in this work. For simplicity,
we show only the results for three distances, namely, $20$, $60$ and $100$ kpc.
For all the distances tested, the \ac{CNN} achieves a higher \ac{TAP} for any
given \acp{FAP} for magnetorotational than neutrino-driven signals.  That is
not surprising as the intrinsic amplitudes for magnetorotational signals are
higher than those of neutrino-driven signals.

\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:ROClog1}
            \includegraphics[width=0.475\textwidth]{ROC_curves_neutrino.png}%
        }\quad
        \subfigure[]{
            \label{fig:ROClog2}
            \includegraphics[width=0.475\textwidth]{ROC_curves_mag.png}%
        }
    \end{center}
    \caption{\ac{ROC} curves showing the classification performance of the
\ac{CNN} for \ac{CCSN} signals of different explosion mechanisms at three
distances, $20$, $60$, and $100$ kpc. The left panel shows the results for the
neutrino-driven mechanism, while the panel on the right shows those for the
magnetorotational mechanism. In both panels, the solid lines are for the
network H+L+VK, while the dashed lines are for HLVK. The black lines in both plots indicate 
the worst scenario where a model has zero ability in distinguishing its input.
\label{fig:ROClog}} 
\end{figure*}

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  In
this figure, a similar trend is seen that magnetorotational signals are easier
for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $0.1\%$, the \ac{CNN} achieves a \ac{TAP} of $82\%$ and $66\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.001\%$, the \ac{CNN} still achieves \acp{TAP} of $72\%$ and $55\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, the networks achieve a
\ac{TAP} of $77\%$ (H+L+VK) or $54\%$ (HLVK) at a \ac{FAP} of $0.1\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $80$ kpc, the \acp{TAP} are no less than $50\%$ 
for H+L+VK for all chosen \acp{FAP} values, and is $66\%$ if
the \ac{FAP} is $0.1\%$. These numbers indicate that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. For sources at $100$, $150$ and $200$ kpc, the \acp{TAP}
drop to $55\%$, $33\%$ and $16\%$ respectively for \ac{FAP} equal to $0.1\%$.
On the other hand, it is more difficult for the
\ac{CNN} to detect and classify the neutrino-driven signals, due to their
weaker amplitudes. For example, for sources at $10$ kpc, the \ac{CNN} achieves
a \ac{TAP} of $38\%$ and $10\%$ for H+L+VK and HLVK respectively if the
\ac{FAP} is $0.1\%$. In Fig.~\ref{fig:eff}, only the values of \ac{TAP} that
are larger than zero will be plotted.

>>>>

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae out to the Large or even Small Magellanic Clouds. 
A neutrino-driven event is likely to be detectable with a network of H+L+VK
and HLVK if the distances is within $5$ kpc and $3$ kpc respectively.
%if the explosion mechanism is the neutrino-driven mechanism.  
By testing the convolutional neural network 
with waveforms not used for training, we show that the true alarm probability
is $66\%$ at $60$ kpc for waveforms $\text{R3E1AC}$ and 
$\text{R4E1FC\_L}$ with a network of H+L+VK.  
For waveforms $\text{s}20$ and $\text{SFHx}$,
the true alarm probabilities are $56\%$ for H+L+VK at $7$ kpc and  $50\%$ for HLVK at $5$ kpc 
respectively.
All at false alarm probability equal to $0.1\%$. 
\end{abstract}

@@

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to
\ac{aLIGO}, \ac{AdVirgo} and KAGRA interferometers. 
It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffen by the strong nuclear force as the core exceeds
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

@@

%
% introduce machine learning
% 
In recent years, the field of machine learning and its sub-field, deep
learning, have been rapidly developing and have shown great potential in many
scientific fields~\cite{krizhevsky2012imagenet, NIPS2014_5423,
simonyan2014very, chen2014semantic, zeiler2014visualizing, szegedy2015going}.
For example, deep learning has been successfully applied to fields including
medical diagnosis~\cite{kononenko2001machine}, object
detection~\cite{redmon2016you}, image
recognition/processing/generation~\cite{he2016deep, krizhevsky2012imagenet,
zhang2016colorful, karpathy2015deep}, and  language
processing~\cite{lample2016neural}. In \ac{GW} astronomy, deep learning has
mostly been applied to the identifications of detector noise artefacts
(glitches)~\cite{mukund2017transient, zevin2017gravity, george2017deep}, the
detection of astrophysical signals~\cite{george2018deep, gabbard2018matching, astone2018new},
and their parameter estimation~\cite{2019arXiv190906296G}.
An advantage of the use of a \ac{CNN} in the detection
of a \ac{GW} signal is that a \ac{CNN} is relatively computationally cheap
compared to other more traditional methods. This is because the heavy
computational work is usually done during the training stage of a \ac{CNN}
prior to its actual application~\cite{goodfellow2016deep}.

%
% the main point of the paper
%
In this work, we demonstrate that a \ac{CNN} can be applied to the detection of
\acp{GW} from \acp{CCSN} and to the classification of their explosion
mechanisms. 
To train our \ac{CNN}, we use simulated \ac{CCSN} waveforms from a
number of studies and simulate sources at a range of distances from $0.2$ to
$200$ kpc for two networks of four \ac{GW} detectors. The first network
consists of \ac{aLIGO}, \ac{AdVirgo} and KAGRA. For the second network, we
still include the detectors of \ac{AdVirgo} and KAGRA, but replace the two
detectors of \ac{aLIGO} with a modest set of planned upgrade version of them -
LIGO A+ in Hanford and Livingston~\cite{miller2015prospects, LIGOW}. All detectors 
are at their design sensitivities. We then
apply the \ac{CNN} to waveforms excluded from the training set to show 
the performance on an independent testing set.
Similar to~\cite{astone2018new}, 
we use machine learning techniques for the search of \acp{GW} from \acp{CCSN}.
However, the \ac{CNN} developed in ~\cite{astone2018new} takes time-frequency images of \ac{GW} detector 
data from the coherent WaveBurst pipeline~\cite{klimenko2005constraint} as input, 
while we aim to develop a \ac{CNN} that works independently and takes time series data from \ac{GW} detectors as input.

@@

%
% describe data pre-processing
%
Since the waveforms are generated at various distances, sampling rates and
durations, it is necessary to normalise the waveforms before they can be used
for the generation of the time series. To do this, we first scale the
amplitudes of the waveforms by moving the sources to $10$ kpc from earth. We
then ensure that the sampling rate are identical for all the waveforms by down
sampling them to a pre-selected sampling rate of $4096$Hz. The longest duration
$\tau$ among the waveforms is then identified and each of the remaining
waveforms is padded with zeros to this duration. To introduce as few artefacts
as possible, a high pass filter with a low cut-off frequency equal to $11$Hz
and a tukey window ($\alpha = 0.08$) are applied prior to the zero padding.

%
% data augmentation
%
To balance the difference in the number of waveforms between the two mechanisms
in the training, validation and testing data set, $36$ exact copies of each neutrino-driven waveform are
made. This will make the ratio of the waveforms modelled by the two mechanisms close to 1, 
while does not change the ratio of the waveforms from different studies modelled by the neutrino-driven mechanism. 
This means each waveform in the neutrino-driven mechanism will have equal representations in the data set,
and waveforms from different mechanisms will have equal representation.
After this procedure, the simulated waveforms are then $\bold{S}(t) = \{\bold{s}_{1}(t), \bold{s}_{2}(t),
..., \bold{s}_{m}(t)\}$, where $m$ is the number of the waveforms and
$\bold{s}_{q}$ is the $q^{\text{th}}$ waveform, defined by, 
%
\begin{equation}\label{eq:ht}
\bold{s}_q(t) = \left(\begin{array}{c} h_q^+(t) \\ h_q^\times(t)
\end{array}\right), 
\end{equation} 
%
where $h_q^+(t)$ and $h_q^\times(t)$ are the two polarisations of the waveform
and $t$ is the time. In this work, we assume a 
fixed GPS time when generating the training/validation/testing data. 
At this GPS time, the value of $t$ is zero.
As mentioned, some of the waveforms used in the work are
generated with the simulations being axisymmetric, in which case the waveforms
are entirely described by one polarisation $h_q^+(t)$. The corresponding
$h_q^\times(t)$ for these waveforms are defined as a vector of zeros. 
%
% rescaling the amplitudes for differnet distances
%
In this work, we perform simulations for distances equal to $0.2$, $0.5$, $1$, $3$, $5$, $7$, $10$, $20$, $30$,
$40$, $50$, $60$, $80$, $100$, $150$ and $200$ kpc. If for a training
procedure, the default distance $d_\text{L}$ of a waveform is not $10$ kpc, the amplitude are rescaled 
by the distance since the amplitude is inversely proportional to the distance.

@@

The next step is to generate simulated time series for the \ac{GW} detectors in

@@

%The input data consist of simulated time series from a network of \ac{GW}

@@

into three exclusive classes, in total three types of time series are

@@

\begin{multline}

@@

\bold{N}_j(T)~~~~~~~~~~~~~~~~~T < T_{\text{max}} p,~T > \tau + T_{\text{max}} p; \\

@@

@@

\end{multline}

@@

For each time series of the background noise class, independently simulated Gaussian

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$310$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 1.26 \times 10^{6}$ data samples, where each of the 3 classes
contains approximately $4.2\times10^{5}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time series for a single detector. A data sample consists of such a
time series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1.0\times10^{4}$ samples being randomly selected for validation.
Of the remainder, $1.70\times10^{5}$ for training and $1.08\times10^{6}$ for tesing. 
The large number of testing samples is chosen so that low values of \ac{FAP} can be obtained 
(see section \ref{sec:result} and section \ref{sec:unseen}).
When the training is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

%
% plot of representative data sample
%
\begin{figure}
\includegraphics[width=0.475\textwidth]{datasample.png}
\caption{A representative example of the simulated time series used to
train/validate/test the \ac{CNN}. The blue shows a whitened time series with a
signal added to Gaussian noise generated using the power spectral density of \ac{aLIGO}. 
The red curve shows the same whitened signal free of
noise. The optimal signal to noise ratio for this signal is $5.5$.
The original waveform is shown in Fig. \ref{fig:neusample1}.
The source is at a distance of $10$ kpc from earth.
\label{fig:sample}} 
\end{figure}

@@

%
% describe ROC curve plots (Fig 5)
%

Although the \ac{CNN} has been trained for all the distances, 
for the neutrino-driven mechanism, only results for distances $\leq 10$ kpc
will be shown. Similarly, only results for distances $\geq 10$ kpc will be displayed 
for the magnetorotational mechanism,. 
This is because the intrinsic amplitudes for magnetorotational signals are
higher than those of neutrino-driven signals and the detectable distances for the two mechanisms 
are generally different.
In Fig.~\ref{fig:ROClog}, we show the \acp{ROC} for both of the \ac{CCSN}
mechanisms and \ac{GW} detector networks tested in this work. For simplicity,
for each mechanism, we show only the results for three distances, 
namely, $0.2$, $1$, $7$ kpc for the neutrino-driven mechanism, and 
$20$, $60$ and $100$ kpc for the magnetorotational mechanism.

\begin{figure*}
     \begin{center}
        \subfigure[the neutrino-driven mechanism]{
            \label{fig:ROClog1}
            \includegraphics[width=0.475\textwidth]{ROC_curves_neutrino.png}%
        }\quad
        \subfigure[the magnetorotational mechanism]{
            \label{fig:ROClog2}
            \includegraphics[width=0.475\textwidth]{ROC_curves_mag.png}%
        }
    \end{center}
    \caption{\ac{ROC} curves showing the classification performance of the
\ac{CNN} for \ac{CCSN} signals of different explosion mechanisms at three
distances, $0.2$, $1$ and $7$ kpc for the neutrino-driven mechanism, and $20$, $60$, 
and $100$ kpc for the magnetorotational mechanism. 
The left panel shows the results for the
neutrino-driven mechanism, while the panel on the right shows those for the
magnetorotational mechanism. In both panels, the solid lines are for the
network H+L+VK, while the dashed lines are for HLVK. The black lines in both plots indicate 
the worst scenario where a model has zero ability in distinguishing its input.
\label{fig:ROClog}} 
\end{figure*}

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  
%In this figure, a similar trend is seen that magnetorotational signals are easier
%for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $0.1\%$, the \ac{CNN} obtains a \ac{TAP} of $82\%$ and $66\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.001\%$, the \ac{CNN} still achieves \acp{TAP} of $72\%$ and $55\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, the networks achieve a
\ac{TAP} of $77\%$ (H+L+VK) or $54\%$ (HLVK) at a \ac{FAP} of $0.1\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $80$ kpc, the \acp{TAP} are no less than $50\%$ 
for H+L+VK for all chosen \acp{FAP} values, and is $66\%$ if
the \ac{FAP} is $0.1\%$. These numbers indicate that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. For sources at $100$, $150$ and $200$ kpc, the \acp{TAP}
drop to $55\%$, $33\%$ and $16\%$ respectively for \ac{FAP} equal to $0.1\%$.
On the other hand, it is more difficult for the CNN to detect and classify the
neutrino-driven signals, due to their weaker amplitudes.
Nonetheless, the \ac{CNN} obtains a \ac{TAP} of $90\%$ and $75\%$ for sources 
at $3.0$ kpc at a \ac{FAP} of $0.1\%$ with
a network of H+L+VK and HLVK respectively.
The \ac{TAP} drops to  $67\%$ for H+L+VK at $5$ kpc.
%it is more difficult for the
%\ac{CNN} to detect and classify the neutrino-driven signals, due to their
%weaker amplitudes. 
%For example, for sources at $10$ kpc, the \ac{CNN} achieves
%a \ac{TAP} of $38\%$ and $10\%$ for H+L+VK and HLVK respectively if the
%\ac{FAP} is $0.1\%$. In Fig.~\ref{fig:eff}, only the values of \ac{TAP} that
%are larger than zero will be plotted.
